\plaintitle

### 2016-11-20 Paper/Rebuttal.txt

Bootstrapped CIs:
Bootstrapped CIs appear frequently in graphical perception work [8, 17, 29]. Prior work ([8], section 4.4) justifies the use of bootstrapping as a robust and non-parametric way of creating CIs.

Midmeans (aka Interquartile Means, or IQMs):
Prior work [8, 17] indeed uses "the means of the midmeans of the log absolute errors" in charts, including Fig. 16 in [8] (which R3 correctly cites as a model for figures in our paper). R3 notes some disadvantages of IQMs: the distribution of such values tends to violate the assumptions of normality of sampling error, which is key for t-tests and ANOVAs. In the paper, we use IQMs only for plots and for reporting effect sizes; for ANOVAs and other inferential tests, we use standard means. We will clarify this distinction in the paper.

While we reiterate that the p-values reported in the paper come from standard means (not IQMs), IQMs have tend to have lower variance than standard means, and so may inflate effect sizes compared with other measures of central tendency. In response to this concern, we generated two-tailed 95% t-confidence intervals of the standard mean. These new CIs do not alter the main results of the paper.

Fig 6 becomes:
sigma, mean(absolute error)+/-95%t-ci
0.05,  0.04 +/- 0.004
0.10,  0.10 +/- 0.005
0.15,  0.15 +/- 0.006
0.20,  0.23 +/- 0.01

Fig 8 becomes:
graph type, mean(error)+/-95%t-ci
area,  -0.01 +/- 0.006
line,  0.01 +/- 0.005
scatterplot, 0.005 +/- 0.003

### 2016-11-21 Paper/Rebuttal.txt

We stand by our decision to present midmeans and bootstrapped confidence intervals. In addition to [8], these methods for showing effect sizes have also been used in more recent graphical perception papers [17, 29]. We will expand our justification for these techniques in the paper, and clarify our methodology.

Bootstrapped CIs:
Bootstrapped CIs appear frequently in graphical perception work [8, 17, 29]. Prior work ([8], section 4.4) justifies the use of bootstrapping as a robust and non-parametric way of creating CIs.

Midmeans (aka Interquartile Means, or IQMs):
Prior work [8, 17] indeed uses "the means of the midmeans of the log absolute errors" in charts, including Fig. 16 in [8] (which R3 correctly cites as a model for figures in our paper). R3 notes some disadvantages of IQMs: the distribution of such values tends to violate the assumptions of normality of sampling error, which is key for t-tests and ANOVAs. In the paper, we use IQMs only for plots and for reporting effect sizes; for ANOVAs and other inferential tests, we use standard means. We will clarify this distinction in the paper.

While we reiterate that the p-values reported in the paper come from standard means (not IQMs), IQMs tend to have lower variance than standard means, and so may inflate effect sizes compared with other measures of central tendency. In response to this concern, we generated two-tailed 95% t-confidence intervals of the standard mean. These new CIs do not alter the main results of the paper.

@@

We will include or clarify study details asked for by reviewers, including the number of exclusions, and re-structuring of the methods section to make the participant information clearer. Though we did submit our study data as a supplement, on acceptance, we would make public our full repository containing data tables, analyses, experimental stimuli, and instruments.

As an alternative study design, both R4 and R5 mention un-constrained input of trend lines (which we had also initially considered). Our design is more concerned with fitting a given model (linear or otherwise) than a freeform "sketch" of a fit, which may or may not correspond to any of the categories of provided models -- and may also suffer from additional noise such as drawing "jitter" or "satisficing" by subjects due to the higher cost of revising drawings. For the targeted research questions in our study, we believe our more parsimonious study design is justified. However, we know of others actively researching the design of experimentally tractable freeform solicitation techniques (e.g., the “draw your guess” NYT example mentioned by R5). We will clarify the distinction between these types of task, citing work in this area that has been published since our submission.

3) Better situate the work in the CHI community:

While many graphical perception works have had a welcome reception at CHI (e.g., [10, 17]), we believe that we can improve the perceived fit of this work at CHI (as per R5) through the inclusion of an implications for design subsection of the conclusion. We believe our results offer guidance to designers in visualization. In particular, they suggest that area charts (a common encoding for time series data) have unanticipated downsides when attempting to convey temporal trends. Similarly, they suggest that standard trend lines (which are OLS lines by default in programs like Excel and other common analysis tools) may not appear credible if there are many outliers in the data; in such cases, viewers may be better off assessing trends visually.
